---
title: "Reproducible Research: Peer Assessment 1"
author: "Alonso Antón"
date: "14/09/2020"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

````{r}

library(data.table)
library(ggplot2)
library(dplyr)

url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

download.file(url,destfile = paste0(getwd(),'/repdata%2Fdata%2Factivity.zip'))

unzip("repdata%2Fdata%2Factivity.zip",exdir="unzip")

````

## Code for reading in the dataset and/or processing the data

```{r}

activity <- data.table::fread("unzip/activity.csv") 

```


## What is mean total number of steps taken per day?

1. Total number of steps by day

```{r}
steps <- activity[,lapply(.SD, sum,na.rm=F),.SDcols=c("steps"),by=c("date")]

head(steps)
```
2. Histogram of the total number of steps taken each day

```{r fig.width=7, fig.height=4}
graphic1 <- steps %>% 
    ggplot(aes(x=steps))+
    geom_histogram(fill="red",binwidth = 1000)+
    labs(title="Daily Steps",x="Steps",y="Frecuency")
print(graphic1)
```

3. Mean and median number of steps taken each day

```{r}
steps_mean <- steps[,lapply(.SD, mean,na.rm=TRUE),.SDcols=c("steps")]

steps_median <- steps[,lapply(.SD, median,na.rm=TRUE),.SDcols=c("steps")]

cbind(steps_mean,steps_median)
```

## What is the average daily activity pattern?

1. Time series plot of the average number of steps taken

```{r}

steps2 <- activity[,lapply(.SD, mean, na.rm=TRUE),.SDcols=c("steps"),by=c("interval")]

steps2 %>% 
    ggplot(aes(x=interval,y=steps))+
    geom_line(color="red",size=1)+
    labs(tittle="Avg. Daily Steps", x= "Interval",
         y="Avg. Steps per day")

```

2. The 5-minute interval that, on average, contains the maximum number of steps

```{r}

steps2[steps==max(steps2$steps),.(max_interval=interval)]

```

## Imputing missing values

1. Identify missing values

```{r}
activity[is.na(steps),.N]
```
2. Replace missing values with median of dataset

```{r}
activity[is.na(steps),"steps"] <- activity[,c(lapply(.SD, median,na.rm=TRUE)),.SDcols=c("steps")]

head(activity)
```

3. Histogram of the total number of steps taken each day after missing values are imputed

```{r}
steps <- activity[,lapply(.SD, sum,na.rm=F),.SDcols=c("steps"),by=c("date")]
steps %>% 
    ggplot(aes(x=steps))+
    geom_histogram(fill="yellow",binwidth = 1000)+
    labs(title="Daily Steps",x="Steps",y="Frecuency")
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
activity <- data.table::fread("unzip/activity.csv")
activity[,date:=as.POSIXct(date,format="%Y-%m-%d")]
activity[,"Day of Week":=weekdays(date)]
activity[grepl(pattern = "lunes|martes|miercoles|jueves|viernes", x = `Day of Week`), "weekday or weekend"] <- "weekday"
activity[grepl(pattern = "sabado|domingo", x = `Day of Week`), "weekday or weekend"] <- "weekend"
activity[, `weekday or weekend` := as.factor(`weekday or weekend`)]
head(activity,10)
```
6. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

````{r}
activity[is.na(steps), "steps"] <- activity[, c(lapply(.SD, median, na.rm = TRUE)), .SDcols = c("steps")]

Interval <- activity[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval, `weekday or weekend`)] 

library(ggplot2)
ggplot(Interval , aes(x = interval , y = steps, color=`weekday or weekend`)) + geom_line() + labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") + facet_wrap(~`weekday or weekend` , ncol = 1, nrow=3)

````